Automated Formal Verification of Stand-alone Solar Photovoltaic Systems

### 2019-07-15 Tex/elsevier/reply Solar Energy JULY 2019.tex
solar energy revised text and reply letter

According to the associate editor and reviewers' comments, our manuscript has been carefully revised and all detailed changes are highlighted in red color, along with the original text. The revision included the Abstract, Introduction, and Results sections.

@@

Note that model checking and software simulations are two different techniques that use mathematical models. However, the way they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then a model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking. This information was included in the revised version of our manuscript in Section X \textcolor{red}{could you please describe the section here?}.

@@

\begin{resposta} 
Indeed, we have missed this particular information in the paragraph that describes the mentioned figure. We thank you so much for this observation. We have revised our manuscript to describe every employed symbol. The revised version of these paragraphs \textcolor{red}{in Section X} are now described as: ``A stand-alone PV system is illustrated in Fig.1. The PV generator is a semiconductor device that can convert solar energy into DC electricity, with high dependence of two weather variables from the site where the system is deployed: solar irradiance $G$ and temperature $T$. For night hours or rainy days, power stored in batteries can be used and it implies the presence of a charge controller. The PV arrays produce DC and therefore when the PV system contains an AC load, a DC/AC conversion is required (inverter). The AC load dictates the behavior of AC electrical load from the house that will be fed by the system. At this pictured modular structure, every element produces/consumes current $I$ and voltage $V$ as illustrated by their physical magnitudes, where $pv$ means photovoltaic, $bat$ is battery, $dc$ is direct and $ac$ is alternating signals.''.
\end{resposta}

@@

\begin{resposta} 
Indeed, we did not inform the impact of changing the computer setup or related issues. We thank the reviewer for this observation. We have revised our manuscript to include the following paragraph: ``Note that the results presented here depend on the computer's processor and memory, the version of each software verifier (i.e., CBMC, ESBMC, CPAchecker), the parameters passed by the command-line, and the algorithm to be solved \textcolor{red}{what do you mean by the algorithm to be solved?}. Additionally, any change from HOMER Pro to another simulation tool can also influence the measured time to obtain results.".
\end{resposta}

@@

The authors hold real load curve and information collected from monitoring system of four (in five) PV systems/cases \textcolor{red}{could you please identify them here? which one does not have such a information?}; however, based on the fact that the automated verification is mathematical and not visual (curves/plots), it seems that our comparison is not an adequate way \textcolor{red}{why not? explain the reason here}.

The information from temperature and solar irradiance from every site, used by the simulation and formal verification tools, come from historical data collected from real stations, which are available in Weather Base and Energy Plus web sites \textcolor{red}{shall we provide the website here?}. Considering that the idea of our study is to show the effectiveness of the proposed verification method just after the design phase (i.e., before deployment), we believe that it is a plausible procedure \textcolor{red}{to....}. Note that the sites from four case studies (i.e., the ones in a riverside indigenous community) do not have weather station available to collect data and the information must be approximated to Manaus data (the closest site with historical data) \textcolor{red}{what are the consequences? shall we state this limitation in the threats to validity section?}.
\end{resposta}

@@

\begin{resposta} 
We thank the reviewer for this comment. We did as suggested, i.e., we changed the section title to ``Verification and Simulation Results" since the results are related to both verification and simulation tools. The data collected from dwellers with real PV systems were used just to confirm flaws and complains about the equipment deployed in the field. The idea was not to present wave forms because how automated verification works is different from that of simulation or testing \textcolor{red}{why is it different? explain here... please do not point to the answer of another reviewer}.
\end{resposta}

@@

We agree with the reviewer that we could further improve the explanation of our study. In our manuscript, we compare model checking with software simulation. However, the way they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then a model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking. This information was included in the revised version of our manuscript in Section X \textcolor{red}{could you please describe the section here?}.

When we work with waveform, we are performing testing of simulation \textcolor{red}{``testing of simulation'' sounds weird. Could you please explain it further here?}, and that is different of automated verification, which is a mathematical proof (of correctness or flaw). Based on the fact that the authors have the five cases deployed in field and with monitoring system in four those PV systems, it is possible to add waveforms to the work. However, the use of this resource seems that is not an adequate way of comparative \textcolor{red}{why not? could you please explain here?}. However, we included details about the estimated load curve of each case study in the revised version of our manuscript (cf. Section $4$), which is represented by a vector with $24$ elements (one to each hour of the day).
\end{resposta}
\end{quote}

>>>>

According to the associate editor and reviewers' comments, our manuscript has been carefully revised and all detailed changes are highlighted in red color, along with the original text. The revision included the Abstract, Introduction, Methodology, and Results sections.

@@

Note that model checking and software simulations are two different techniques that use mathematical models. However, the way they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then a model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking. This information was included in the revised version of our manuscript in \textcolor{red}{Section 1 (Introduction).}

@@

\begin{resposta} 
Indeed, we have missed this particular information in the paragraph that describes the mentioned figure. We thank you so much for this observation. We have revised our manuscript to describe every employed symbol. The revised version of these paragraphs \textcolor{red}{in Section "2.2 Component models for a stand-alone PV system"} are now described as: ``A stand-alone PV system is illustrated in Fig.1. The PV generator is a semiconductor device that can convert solar energy into DC electricity, with high dependence of two weather variables from the site where the system is deployed: solar irradiance $G$ and temperature $T$. For night hours or rainy days, power stored in batteries can be used and it implies the presence of a charge controller. The PV arrays produce DC and therefore when the PV system contains an AC load, a DC/AC conversion is required (inverter). The AC load dictates the behavior of AC electrical load from the house that will be fed by the system. At this pictured modular structure, every element produces/consumes current $I$ and voltage $V$ as illustrated by their physical magnitudes, where $pv$ means photovoltaic, $bat$ is battery, $dc$ is direct and $ac$ is alternating signals.''.
\end{resposta}

@@

\begin{resposta} 
Indeed, we did not inform the impact of changing the computer setup or related issues. We thank the reviewer for this observation. We have revised our manuscript to include the following paragraph in Section "5.2 Objectives and Setup": "Note that the results presented here depend on the computer's processor and memory, the version of each software verifier (i.e., CBMC, ESBMC, CPAchecker), the parameters passed by the command-line, and the \textcolor{red}{coded problem (software created by authors) to be solved. Additionally, any change from HOMER Pro to another simulation tool can also influence the measured time to obtain results}.".
\end{resposta}

@@

The authors hold real load curve and information collected from monitoring system using Raspberri Pi deployed  \textcolor{red}{at House 1, House 2, House 3 and House 4}; however, based on the fact that the automated verification is mathematical and not visual,  \textcolor{red}{(automated verifiers do not present curves or plots; they present the result of FAIL or SUCCESS analysis of coded system, with the counterexample detailing the status of the variables that cause a FAIL in the analysis), so it seems that this comparison is not adequate}.

The information from temperature and solar irradiance from every house location, used by the simulation and formal verification tools, come from historical data collected from real stations, which are available in Weather Base and Energy Plus web sites, \textcolor{red}{as presented at Section "References"}. Considering that the idea of our study is to show the effectiveness of the proposed verification method just after the design phase (i.e., before deployment), we believe that it is a plausible procedure \textcolor{red}{and not using weather base stations in each place. Even because the fact that House 1, House 2, House 3, and House 4 (the ones in a riverside indigenous community) do not have weather station available to collect data and the information must be approximated to Manaus data (the closest municipality with historical data and regular measurement of data). Because of that, we decide to include this issue at the Section "5.4 Threats to Validity". We thank you for that specifically comment}.
\end{resposta}

@@

\begin{resposta} 
We thank the reviewer for this comment. We did as suggested, i.e., we changed the section title to ``Verification and Simulation Results" since the results are related to both verification and simulation tools. The data collected from dwellers with real PV systems were used just to confirm (or not) flaws and complains about the equipment deployed in the field.

The authors hold real load curve and information collected from monitoring system using Raspberri Pi deployed  \textcolor{red}{at House 1, House 2, House 3 and House 4}; however, based on the fact that the automated verification is mathematical and not visual,  \textcolor{red}{(automated verifiers do not present curves or plots; they present the result of FAIL or SUCCESS analysis of coded system, with the counterexample detailing the status of the variables that cause a FAIL in the analysis, as described in Section 4 of the paper), so it seems that this comparison is not adequate}.

Note that model checking and software simulations are two different techniques that use mathematical models. However, the way they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then a model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking. This information was included in the revised version of our manuscript in \textcolor{red}{Section 1 (Introduction).}
\end{resposta}

@@

We agree with the reviewer that we could further improve the explanation of our study. In our manuscript, we compare model checking with software simulation. However, the way they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then a model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking. This information was included in the revised version of our manuscript in Section \textcolor{red}{"1 Introduction"}.

When we work with waveform, we are performing \textcolor{red}{testing or simulation}, and that is different of automated verification. \textcolor{red}{We consider that this demand (number 7 in this Reply Letter) is related to the experimental question done by the same reviewer at item number (6). Therefore we expect that the explanations done at item 6, with the consequent updated paper can cover any further issue about it}. 
\end{resposta}
\end{quote}

### 2019-07-15 Tex/elsevier/trindade_solar_energy_rev1.tex
solar energy revised text and reply letter

\begin{abstract}
With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. Automated verification using model checking has proven to be an effective technique to validate \textcolor{red}{complex systems}. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system including solar panel, charge controller, battery, inverter, and electric load. \textcolor{red}{Therefore the focus is in the solar photovoltaic project validation, as intended to be done just after the system sizing, prior to buy equipment and deploy it at the field, as a way to ensure the intended behavior. Case studies, ranging from 700 W to 1,200 W}, were used to evaluate this proposed approach and to compare that with specialized simulation tool. Different verification tools were evaluated as well, in order to compare performance and soundness among automated tools. \textcolor{red}{The results reported by the automated verification and by the simulation tools were compared with data collected from dwellers of the deployed cases, showing the effectiveness of our approach, where specific conditions that lead to failures in a photovoltaic solar system are only detailed by the automated verification method.}
\end{abstract}

@@

\textcolor{red}{Here we compare model checking with simulation tool. However how they work is completely different. Simulation depends on the choice of input variables (and their values) in order to obtain the output. In the other hand, model checking performs an automatic output search that causes the system to fails (a not expected behavior of the system) and presents the input that causes it. Therefore only with model checking we can prove the absence of system flaws.}

@@

Line 2 represents all the information that comes from the PV sizing and from the equipment manufacturers data: specification and data from PV, batteries, inverter and charge controller. This item includes as information from the house's load curve.

\textcolor{red}{Those estimated load curves are converted as vectors with 24 positions, one to each house of the day (in Watts): 
House 1: [118, 118, 118, 46, 46, 46, 95, 95, 170, 170, 296, 242, 242, 95, 95, 95, 95, 95, 342, 288, 288, 288, 288, 118]; 
House 2: [136, 136, 136, 136, 136, 136, 67, 67, 184, 184, 184, 184, 184, 67, 67, 67, 67, 67, 253, 253, 253, 253, 253, 136]; 
House 3: [113, 113, 113, 113, 113, 113, 67, 67, 217, 97, 97, 97, 97, 97, 97, 97, 97, 97, 263, 113, 113, 113, 113, 113]; 
House 4: [207, 207, 207, 135, 135, 135, 66, 66, 161, 161, 233, 253, 248, 66, 66, 66, 66, 66, 302, 317, 322, 302, 302, 207]; 
House 5: [45, 16, 16, 16, 16, 16, 0, 0, 0, 72, 72, 222, 150, 150, 0, 0, 72, 72, 814, 814, 814, 742, 742, 16].}

@@

\textcolor{red}{Note that the results presented here depend on the computer's processor and memory, the version of each tool (CBMC, ESBMC, CPAchecker), the parameters passed by the command-line, and the algorithm to be solved. Additionally, any change from HOMER Pro to another simulation tool can influence the measured time to obtain results.}

@@

%At the terminal, the command used to perform the verification is:
%\$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-no-div-by-zero-check -\phantom{}-unwind 300 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard --z3
%
%Where:
%
%\begin{itemize}
%\item The first three parameters, after the filename, are related to options that are usual to find bug in software, %for example, like bound check to arrays, pointer check, and division by zero, 
%but unnecessary to check at this kind of problem (if not removed, there is lost of performance during the automated verification);
%\item The parameter $unmind$ tells to ESBMC the limit to unroll the loops. This number was optimized (empirically) in order to reduce the running time and avoid to unwind unnecessarily the loops;
%\item The two parameters with $symex$ tell to ESBMC to perform an incremental SMT solving. There are other options, but this parameter is necessary because the complexity of the algorithms. The incremental SMT solving uses few RAM memory, compared with other SMT solving. 
%During empirical tests of the algorithms, the incremental solving was the only one who do not demanded 100\% the RAM memory. The use of swap-memory, i.e., the use of hard disk, reduces the performance and must be avoided;
%\item And the last parameter says to the tool that the Z3 SMT solver will be used.
%\end{itemize}
%
%\subsection{Experimental results}
%\label{sec:results}
%---------------------------------------------------
\subsection{Threats to Validity}
%---------------------------------------------------
We have reported a favorable assessment of the proposed method. % over a diverse set of real-world benchmarks. 
Nevertheless, we have also identified \textcolor{red}{four} threats to the validity of our results that can further be assessed.

@@

@@

We have considered five case studies, ranging from $700$ W to $1,200$ W; three state-of-art verification engines were considered (ESBMC, CBMC, and CPAchecker); and one specialized simulation tool (HOMER Pro). The results from automated verification and simulation tools were confronted with information collected from dwelles of the sites where the real PV systems were deployed.} Although the verification method proposed takes longer than simulation methods, it is able to present details that lead to failures in a PV system that is not a feature presented in commercial simulation tool. In particular, the proposed method was successful in finding sizing errors and indicating in details. Related to the verification engines comparative, the ESBMC with the Z3 solver executed in the incremental configuration presented the better performance (around four times faster than CBMC), used less RAM memory (less than 2 GB when compared to 9.2 GB of CBMC and 19.2 GB of CPAchecker), and all the results were sound because the PV owners and the monitoring system validated the possible flaws that the system could be presenting in the field.

@@

As future work, we will also consider other types of renewable energy and even hybrid ones to allow our method to verify typical rural electrification.

>>>>

\begin{abstract}
With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. Automated verification using model checking has proven to be an effective technique to validate \textcolor{red}{complex systems}. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system including solar panel, charge controller, battery, inverter, and electric load. \textcolor{red}{Therefore the focus is in the solar photovoltaic project validation, as intended to be done just after the system sizing, prior to buy equipment and deploy it at the field, as a way to ensure the intended behavior. Case studies, ranging from 253 W to 814 W}, were used to evaluate this proposed approach and to compare that with specialized simulation tool. Different verification tools were evaluated as well, in order to compare performance and soundness among automated tools. \textcolor{red}{The results reported by the automated verification and by the simulation tools were compared with data collected from dwellers of the deployed cases, showing the effectiveness of our approach, where specific conditions that lead to failures in a photovoltaic solar system are only detailed by the automated verification method.}
\end{abstract}

@@

\textcolor{red}{Here we compare model checking with simulation tool. However how they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then the model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking.}

@@

Line 2 represents all the information that comes from the PV sizing and from the equipment manufacturers data: specification and data from PV, batteries, inverter and charge controller. This item includes as well information from the house's load curve.

\textcolor{red}{At this work, we estimated load curves based on survey and visiting performed in June of 2017, and the the estimated load curve was converted as vectors with 24 positions, one to each house of the day (in Watts): 
House 1: [118, 118, 118, 46, 46, 46, 95, 95, 170, 170, 296, 242, 242, 95, 95, 95, 95, 95, 342, 288, 288, 288, 288, 118]; 
House 2: [136, 136, 136, 136, 136, 136, 67, 67, 184, 184, 184, 184, 184, 67, 67, 67, 67, 67, 253, 253, 253, 253, 253, 136]; 
House 3: [113, 113, 113, 113, 113, 113, 67, 67, 217, 97, 97, 97, 97, 97, 97, 97, 97, 97, 263, 113, 113, 113, 113, 113]; 
House 4: [207, 207, 207, 135, 135, 135, 66, 66, 161, 161, 233, 253, 248, 66, 66, 66, 66, 66, 302, 317, 322, 302, 302, 207]; 
House 5: [45, 16, 16, 16, 16, 16, 0, 0, 0, 72, 72, 222, 150, 150, 0, 0, 72, 72, 814, 814, 814, 742, 742, 16].}

@@

\textcolor{red}{Note that the results presented here depend on the computer's processor and memory, the version of each software verifier (i.e., CBMC, ESBMC, CPAchecker), the parameters passed by the command-line, and the coded problem (software created by authors) to be solved. Additionally, any change from HOMER Pro to another simulation tool can also influence the measured time to obtain results.}

@@

%At the terminal, the command used to perform the verification is:
%\$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-no-div-by-zero-check -\phantom{}-unwind 300 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard --z3
%
%Where:
%
%\begin{itemize}
%\item The first three parameters, after the filename, are related to options that are usual to find bug in software, %for example, like bound check to arrays, pointer check, and division by zero, 
%but unnecessary to check at this kind of problem (if not removed, there is lost of performance during the automated verification);
%\item The parameter $unmind$ tells to ESBMC the limit to unroll the loops. This number was optimized (empirically) in order to reduce the running time and avoid to unwind unnecessarily the loops;
%\item The two parameters with $symex$ tell to ESBMC to perform an incremental SMT solving. There are other options, but this parameter is necessary because the complexity of the algorithms. The incremental SMT solving uses few RAM memory, compared with other SMT solving. 
%During empirical tests of the algorithms, the incremental solving was the only one who do not demanded 100\% the RAM memory. The use of swap-memory, i.e., the use of hard disk, reduces the performance and must be avoided;
%\item And the last parameter says to the tool that the Z3 SMT solver will be used.
%\end{itemize}
%
%\subsection{Experimental results}
%\label{sec:results}
%---------------------------------------------------
\subsection{Threats to Validity}
%---------------------------------------------------
We have reported a favorable assessment of the proposed method. % over a diverse set of real-world benchmarks. 
Nevertheless, we have also identified \textcolor{red}{five} threats to the validity of our results that can further be assessed.

@@

\textcolor{red}{
\textit{Temperature and Solar Irradiance Data}: Information related to temperature and solar irradiance of each case study, independently of using simulation or formal verification, come from data bases available online. However, considering that riverside communities do not have weather stations, the data used at this work come from the closest municipality (Manaus in all case studies), where stations collect regularly those data. Therefore, the most accurate should be the use of weather stations in each location to obtain the data.}

@@

We have considered five case studies, ranging from \textcolor{red}{$253$ W to $814$ W}; three state-of-art verification engines were considered (ESBMC, CBMC, and CPAchecker); and one specialized simulation tool (HOMER Pro). The results from automated verification and simulation tools were confronted with information collected from dwelles of the sites where the real PV systems were deployed.} Although the verification method proposed takes longer than simulation methods, it is able to present details that lead to failures in a PV system that is not a feature presented in commercial simulation tool. In particular, the proposed method was successful in finding sizing errors and indicating in details. Related to the verification engines comparative, the ESBMC with the Z3 solver executed in the incremental configuration presented the better performance (around four times faster than CBMC), used less RAM memory (less than 2 GB when compared to 9.2 GB of CBMC and 19.2 GB of CPAchecker), and all the results were sound because the PV owners and the monitoring system validated the possible flaws that the system could be presenting in the field.

@@

As future work, we will \textcolor{red} {expand the number of case studies (preferentially in different countries), and also consider} other types of renewable energy and even hybrid ones to allow our method to verify typical rural electrification.
